---
title: "README"
output: html_document
---


## sshd

vim /etc/ssh/sshd_config
And be sure this options are set:

ClientAliveInterval 60
ClientAliveCountMax 0


## Setup

```{sh}
# rgl<-mixOmics
apt install xorg libx11-dev libglu1-mesa-dev
# gsl <- userfrienndly
apt install libgsl-dev libxml2-dev

## animation requires imagemagick
sudo apt install libmagick++-6.q16-dev

## fonts
add-apt-repository multiverse
apt update
apt install ttf-mscorefonts-installer

## metaboanalystr
apt install libnetcdf-dev
```

## R packages

- mixOmics: spls.
- multcompView: Posthoc-test significance label.
- fs: path_sanitize

OpenMx compilation 

```
# sudo vim /opt/microsoft/ropen/3.5.3/lib64/R/etc/Makeconf
CXX14 = g++
CXX14FLAGS = -DU_STATIC_IMPLEMENTATION -g -O2
CXX14PICFLAGS = -fpic
CXX14STD = -std=gnu++14
```


```{r}

pkgs <- c(
    'fs', 'pdftools', 'tiff', 'readr',
    'extrafont',
    'FactoMineR', 'factoextra', 'broom', 'data.table', 'dunn.test',
    'DT', 
    'dplyr', 'ggpubr', 'ggrepel', 'ggplot2', 'ggsignif', 'animation',
    'gridExtra', 'igraph', 'knitr', 'magrittr', 'Hmisc',
    'markdown', 'rmarkdown', 'rprojroot', 'multcompView', 'pander', 'pheatmap',
    'shiny', 'tableone', 'testthat', 'tidyr', 'userfriendlyscience', 'FSA',
    'caret', 'pROC',
    'BiocManager')

pkgs <- pkgs[!pkgs %in% installed.packages()]
install.packages(pkgs, repos = 'https://mirrors.tuna.tsinghua.edu.cn/CRAN')
```

```{shell}
R CMD config CXX11
echo $(R RHOME)/etc/Makeconf
```

## Bioconductor mirror

```{r, eval=F, echo=T}
# options(BioC_mirror="https://mirrors.tuna.tsinghua.edu.cn/bioconductor")
chooseBioCmirror()
BiocManager::install(c('mixOmics', 'ropls'))

library(extrafont)
font_import()
```



## Font

- Add repository for ttf-mscorefonts-installer 
- Install ttf-mscorefonts-installer 
- Import to R using extrafont

## Administration

Disconnected from the server
https://docs.rstudio.com/shiny-server/

https://stackoverflow.com/questions/39116782/disconnected-from-server-in-shinyapps-but-locals-working
https://docs.rstudio.com/shiny-server/#sockjs_heartbeat_delay

```
Increase the value of 'sockjs_heartbeat_delay' in shiny-server.conf 
```

https://github.com/rstudio/shiny-server/pull/264

```
http_keepalive_timeout 220;
sockjs_disconnect_delay 200;
disable_websockets;
```

## Build a dynamic UI that reacts to user input

http://shiny.rstudio-staging.com/articles/dynamic-ui.html#addingremoving-ui-with-insertui-and-removeui


## options("repos") only changed to MRAN when scanForPackages = TRUE and there is a new package to install

https://github.com/RevolutionAnalytics/checkpoint/issues/274

```{r, eval=F}
checkpoint::checkpoint("2019-03-13", scanForPackages = TRUE)

# To change the CRAN mirror to MRAN mirror of specified date
checkpoint::setSnapshot("2019-03-13")

# Make sure that packages are loaded from checkpoint directory
library(data.table, lib.loc = .libPaths()[1])
```


